---
title: "PS270 Final Project"
author: "Grace Fragakis"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to fashion, or environmental issues. I would consider combining the two and finding data related to environmental damages in different countries due to fashion trends or consumerism. 

I am also interested in exploring data related to the amount of soda consumed in each state of the United states, specifically Diet Dr. Pepper. 

Furthermore, I am also interested in exploring data related to the size of celebrities carbon footprint compared to the average human in the US. 

## Proposal

For my research question I will be exploring: How does increasing atmospheric carbon dioxide concentration in one area with an extreme physical stature, such as Mauna Loa, compare to the rest of the world? Carbon dioxide emissions are most prominently caused by us humans, and our consumerism. Between traveling, buying, and using, all of our carbon footprints put our planet in danger. I believe that carbon dioxide concentrations in our atmosphere in an extreme environment will be similar to the rest of the world despite the difference in geographic makeup. Therefore, if carbon emissions are increased in one area, then they will also increase in another area.  My explanatory variable that I would measure is the carbon dioxide concentration in the atmosphere in Mauna Loa, which is measured by parts per million. My outcome variable of interest is the comparison of carbon dioxide in a different part of the world, also measured in parts per million. The observed pattern in data that would support my hypothesis are similar carbon dioxide levels in Mauna Loa compared to the rest of the world.To disprove my hypothesis extreme differences in the atmospheric mean of carbon dioxide would need to be recorded. I will be using the data set: CO2 PPM Trends in Atmospheric Carbon Dioxide


```{r fig.width=10, fig.height=6}
library(tidyverse)

gl_mean <- read_csv('data/co2-annmean-gl.csv') |>
  mutate(Source = "Global Annual Mean")
mlo_mean <- read_csv('data/co2-annmean-mlo.csv')|>
  mutate(Source = "Mauna Loa")

combined_data <- bind_rows(gl_mean, mlo_mean) |>
  filter(Year >= 1979)

ggplot(combined_data, aes(x = Year, y = Mean, color = Source)) +
  geom_line() +
  labs(title = "Comparison of CO2 Levels (Global Annual Mean vs. Mauna Loa)",
       x = "Year",
       y = "Mean CO2 Concentration (ppm)") +
  scale_color_manual(values = c("Global Annual Mean" = "blue", "Mauna Loa" = "red"))

```

```{r}

ggplot(gl_mean, aes(x = Mean)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density Plot of CO2 Concentrations (Global Annual Mean)",
       x = "Mean CO2 Concentration (ppm)",
       y = "Density")


```

Based on these two graphs I have created from my selected data set it seems as though my hypothesis is beginning to be proven. It appears that the carbon dioxide levels in an extreme location, such as Mauna Loa, seems to be aligning with the global annual mean of carbon dioxide levels throughout the rest of the world. This is interesting because Mauna Loa is the largest active volcano on Earth, which takes up over half the island of Hawai'i. The second graph shows how the global annual mean of the density of carbon dioxide concentrations is becoming less dense yet is approaching a higher level of parts per million detected. 
