---
title: "PS270 Final Project"
author: "Grace Fragakis"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---




## Introduction

For my research question I will be exploring: How does increasing atmospheric carbon dioxide concentration in one area with an extreme physical stature, such as Mauna Loa, compare to the rest of the world over time? The null hypothesis is: There is no significant difference in the rate of change of atmospheric carbon dioxide concentration at Mauna Loa compared to the rest of the world over time. The alternate hypothesis is: The rate of change of atmospheric carbon dioxide concentration at Mauna Loa is significantly different from that of the rest of the world over time. This research question is important and interesting to me because carbon dioxide emissions are most prominently caused by us humans, and our consumerism. Between traveling, buying, and using, all of our carbon footprints put our planet in danger. I believe that carbon dioxide concentrations in our atmosphere in an extreme environment will be similar to the rest of the world despite the difference in geographic makeup. Therefore, my proposed hypothesis is if carbon emissions are increased in one area, then they will also increase in another area. My explanatory variable that I would measure is the carbon dioxide concentration in the atmosphere in Mauna Loa, which is measured by parts per million. My outcome variable of interest is the comparison of carbon dioxide for the global annual mean, also measured in parts per million. The observed pattern in data that would support my hypothesis are similar carbon dioxide levels in Mauna Loa compared to the rest of the world.To disprove my hypothesis extreme differences in the atmospheric mean of carbon dioxide would need to be recorded. I will be using the data set: CO2 PPM Trends in Atmospheric Carbon Dioxide.


## Data Section

For data analysis I will be using a cross-sectional research design. I will be observing the changes in carbon dioxide in the environment over a period of time–1980 to present day. The source of data I am using to prove my hypothesis is CO2 PPM Trends in atmospheric Carbon Dioxide, which closely observes the changes in atmospheric carbon dioxide over time. The key dependent and independent variables of interest will be measured using both a statistical model and expert coding. I chose to use a statistical model because this model is expressing the relationship between the independent and dependent variables I am interested in as a function. 
It is essential for my research inquiry to see a trending line over time for both Mauna Loa and the average global CO2 mean. The plots I have created below are the exact data that I need in order to further observe what I have intended to answer. Starting in 1980, this graph allows me to perceive the changes in carbon dioxide closely. Having a wide spread timeline is crucial for this study because as the United States and the world have developed and evolved there have been new innovations that will impact the spread of the data. The dependent variable that I have chosen will be easily determined to do what I hypothesized using this plot with these specific variables and parameters. 




```{r fig.width=10, fig.height=6}
library(tidyverse)
library(ggplot2)

gl_mean <- read_csv('data/co2-annmean-gl.csv') |>
  mutate(Source = "Global Annual Mean")
mlo_mean <- read_csv('data/co2-annmean-mlo.csv')|>
  mutate(Source = "Mauna Loa")

combined_data <- bind_rows(gl_mean, mlo_mean) |>
  filter(Year >= 1979)

ggplot(combined_data, aes(x = Year, y = Mean, color = Source)) +
  geom_line() +
  labs(title = "Comparison of CO2 Levels (Global Annual Mean vs. Mauna Loa)",
       x = "Year",
       y = "Mean CO2 Concentration (ppm)") +
  scale_color_manual(values = c("Global Annual Mean" = "blue", "Mauna Loa" = "red"))

```

```{r}

         
ggplot(gl_mean, aes(x = Mean)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density Plot of CO2 Concentrations (Global Annual Mean)",
       x = "Mean CO2 Concentration (ppm)",
       y = "Density")


```


Based on these two graphs I have created from my selected data set it seems as though my hypothesis is beginning to be proven. It appears that the carbon dioxide levels in an extreme location, such as Mauna Loa, seems to be aligning with the global annual mean of carbon dioxide levels throughout the rest of the world. This is interesting because Mauna Loa is the largest active volcano on Earth, which takes up over half the island of Hawai'i. The second graph shows how the global annual mean of the density of carbon dioxide concentrations is becoming less dense yet is approaching a higher level of parts per million detected. 

```{r}
         
data_w_mean <- combined_data |>
  group_by(Source) |>
  summarize(overall_mean_CO2 = mean(Mean, na.rm = TRUE))

ggplot(data_w_mean, aes(x = Source, y = overall_mean_CO2, fill = Source)) +
  geom_col() +
  labs(title = "Comparison of CO2 Levels (Global Annual Mean vs. Mauna Loa)",
       x = "CO2 Source",
       y = "Mean CO2 Concentration (ppm)") +
  scale_fill_manual(values = c("Global Annual Mean" = "blue", "Mauna Loa" = "red"))

library(infer)
set.seed(02138)

p_val <- combined_data |>
  specify(Mean ~ Source) |>
  generate(reps = 1000, type = "bootstrap") |>
  calculate(stat = "diff in means", order = c("Global Annual Mean", "Mauna Loa")) |>
  get_p_value(
    obs_stat = data_w_mean$overall_mean_CO2[1] - data_w_mean$overall_mean_CO2[2], 
    direction = "two-sided" 
  )

p_val

```

## Results

The plots that I have created prove my hypothesis to be true. The first plot I created showing the two trajectories of Mauna Loa and the global annual mean of carbon dioxide levels shows a close relationship that causes me to infer that even when an extreme location such as Mauna Loa has changes in carbon dioxide levels, that the rest of the world also experiences these changes as well. The main coefficient of interest for this research plot was the carbon dioxide levels of an extreme location that is geographically different than most areas. As I review how this plot has displayed the concentrations of carbon dioxide, I conclude that for every ten years that carbon dioxide concentrations have been observed, not only are the average concentrations increasing for Mauna Loa, but they also are increasing for the rest of the world. These results are statistically significant because the p-value is larger than 0.05, therefore we fail to reject the null hypothesis. I believe that they should be interpreted causally, not only does the cause happen before the effect, but there is a close relationship between the two variables. 
The second plot I have created measures the density of carbon dioxide concentrations for the global annual mean. It was essential that the density was measured because we need to understand the extent of human impact on the climate and understand how extreme the concentrations of carbon dioxide are for the world. This plot shows that as the mean carbon dioxide concentrations increase, the density of the actual concentrations decreases. The main coefficient of interest for this plot was the density levels and this relationship between the actual concentration of carbon dioxide and the density level is very interesting. These results should be interpreted causally because like the first graph that I created, the cause happens before the effect–concentration before density–and there is a definite relationship between the two variables. 
The final graph I created is similar to the first graph, however it is a closer side by side comparison and shows how close the two means are between Mauna Loa and the rest of the globe, which is exactly what I was looking for in this research project. This graph clearly shows that carbon dioxide emissions in an extreme location is very similar to the rest of the world, regardless of confounding variables. Like the first graph, the main coefficient of interest for this research plot was the carbon dioxide levels of an extreme location that is geographically different from most areas. Both locations were almost measured at 400 ppm. The results however, were not found to be statistically significant because the p-value was greater than 0.05, so we fail to reject the null hypothesis of this research. Even though we cannot reject the null hypothesis of this research, that means that the difference in these two variables were so close which was what I was attempting to prove through my research. I believe these results should be interpreted casually, however, we must be aware of confounding variables as this research is observational. 


## Conclusion

After observing the carbon dioxide levels in Mauna Loa and the rest of the world’s average I can confidently conclude my hypothesis has been proven. It is true that when carbon dioxide levels increase in a geographically different location than most of the world, that carbon dioxide levels still do increase in a very similar way for the rest of the globe. Even though the results were nast statistically significant, the amounts of CO2 for both locations were extremely close, which is what I aimed to prove. Limitations of this research include confounding variables, such as determining the difference between natural releases of carbon dioxide on Mauna Loa since it is an active volcano and actual emissions of carbon dioxide. More confounding variables are local weather patterns, altitude of Mauna Loa, and local land use of the areas. There is also missing data in this experiment because it would help my study to determine the difference in density in Mauna Loa and the rest of the world and how the actual concentrations compare like I did for my second plot.  If I had more time and money to conduct this research I would investigate how these confounding variables could be eliminated. This could be done by studying directly in Mauna Loa and researching more of their local land use and monthly activity of the volcano. I think by doing this my research could become more exact and I could closely follow how these factors influence the concentrations of carbon dioxide on Mauna Loa. 
